%%
%% Author: lukas
%% 03.01.2019
%%

\section{Algorithms}\label{sec:algorithms}

This work does not contain any own algorithm implementation for scheduling problems,
instead I would like to use pre-prepared and already implemented solver.
First we must specify which kind of approach we would like to choose.
We have many options, how to represent and then solve scheduling problem such as
\begin{itemize}
    \item Linear programming
    \begin{itemize}
        \item more specifically mixed integer programming
    \end{itemize}
    \item Constraint programming
    \begin{itemize}
        \item more specifically heuristics algorithms
    \end{itemize}
\end{itemize}

\subsection{Linear Programming}\label{subsec:linear-programming}
Some general information about linear programming goes here %TODO add here some info

\subsubsection{Advantages of linear programming approach}
%TODO

\subsubsection{Disadvantages of linear programming approach}
%TODO


\subsection{Heuristic algorithms}\label{subsec:heuristic-algorithms}
General information about heuristic %TODO add heuristic

\subsubsection{Advantages of heuristic algorithms}
The main advantage of heuristic algorithms is that they offer a quick solution for problem they are solving.

\subsubsection{Disadvantages of heuristic algorithms}
The main downside of HA is the fact, that they can't guarantee that found solution is the optimal one.

\subsection{Existing solutions}\label{subsec:existing-solutions}
After considering all advantages and disadvantages of previously mentioned approaches, I decided to use \textit{heuristic algorithm}
in my implementation of RMS\@.
Since it is not main goal of this work to implement such a algorithm I will choose some existing implementation of some heuristic algorithm which
is general enough to be used in my paper.\\
I would like to present two selected existing implementations
\begin{itemize}
    \item OptaPlanner
    \item TASP
\end{itemize}

\subsubsection{OptaPlanner}
OptaPlanner is generic heuristics based constraint solver.
%TODO more fom https://www.optaplanner.org

\subsubsection{TASP}
\textit{Task and Asset Scheduling Platform}  is a lightweight framework developed by Blindspot Solutions designed to solve a large
variety of optimization and scheduling problems from the area of logistics, workforce management, manufacturing, planning and others.
It contains a modular, efficient planning engine utilizing latest optimization algorithms.
TASP is delivered as a software library to be used through its API in applications which require powerful scheduling capabilities.

\subsection{Selected algorithm}\label{subsec:selected-algorithm2}
I decided to use TASP in for my paper, because \ldots %TODO add explanation why I selected TASP

