\documentclass[thesis=B,english]{BPthesis}[2012/06/26]

% basic packages
\usepackage[utf8]{inputenc} % input in utf8
\usepackage{cite} % bibtex
\bibliographystyle{ieeetr}

\usepackage{url} % so URLs look better in bib
\usepackage{amsmath} % math extensions

% fancy packages
\usepackage{indentfirst} % indent every time after new line
\usepackage{xcolor} % various colors used in listings and in template

\hypersetup{hidelinks=true} % disable hideous link borders

% specific packages
\usepackage[linesnumbered,boxed]{algorithm2e} % algorithm engine, lines are numbered and it is in box
\SetKwInput{Input}{Input} % set keyword for algorithm2e
\SetKwInput{Output}{Output} % set keyword for algorithm2e
\SetKwInput{Result}{Result} % set keyword for algorithm2e

\usepackage{listings} % programming examples
\lstdefinelanguage{Kotlin}{ % Kotlin code in the document
  comment=[l]{//},
  commentstyle={\color{gray}\ttfamily},
  emph={delegate, filter, first, firstOrNull, forEach, lazy, map, mapNotNull, println, return@},
  emphstyle={\color{black}},
  identifierstyle=\color{black},
  keywords={abstract, actual, as, as?, break, by, class, companion, continue, data, do, dynamic, else, enum, expect, false, final, for, fun, get, if, import, in, interface, internal, is, null, object, override, package, private, public, return, set, super, suspend, this, throw, true, try, typealias, val, var, vararg, when, where, while},
  keywordstyle={\color[RGB]{0 0 128}\bfseries},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
  morestring=[s]{"""*}{*"""},
  ndkeywords={@Route, @Deprecated, @JvmField, @JvmName, @JvmOverloads, @JvmStatic, @JvmSynthetic, Array, Byte, Double, Float, Int, Integer, Iterable, Long, Runnable, Short, String},
  ndkeywordstyle={\color[RGB]{0 0 128}\bfseries},
  sensitive=true,
  stringstyle={\color[RGB]{0 128 0}\ttfamily},
  backgroundcolor=\color[RGB]{244 244 244}
}

\newcommand{\inlinecode}[1]{\colorbox[RGB]{244,244,244}{\lstinline[language=Kotlin]$#1$}} % one line kotlin code
\newcommand{\inlinedata}[1]{\colorbox[RGB]{244,244,244}{\texttt{#1}}} % one line of data

% TODO remove when ready
\usepackage[disable]{todonotes} % don't generate todos to pdf
% \usepackage{todonotes} % show todos in generated pdf

%template setup
\department{Department of Computer Science}
\title{Design and implementation of a scalable server for parallelization of optimization algorithms execution}
\authorGN{Lukáš}
\authorFN{Forst} 
\authorWithDegrees{Lukáš Forst}
\author{Lukáš Forst}
\supervisor{Ing. Ondřej Vaněk, Ph.D.}
\placeForDeclarationOfAuthenticity{Prague}
\declarationOfAuthenticityOption{1} % 1-6

\acknowledgements{
  \noindent
  I would first like to thank my thesis advisor, Ondřej Vaněk,
  whose office's door was always open whenever I ran into a trouble spot or had a question about my research or writing. 
  
  \noindent
  I would like to also thank my colleagues in Blindspot Solutions and especially to Petr Eichler,
  whose valuable comment suggestions on my paper inspired me to improve the quality of the assignment
  and helped me to overcome the various challenges I faced during the development.
  
  \noindent
  Finally, I must express my very profound gratitude to my parents and to my better half for providing me with unfailing support and continuous encouragement throughout my years of study and through the process of researching and writing this thesis. 

  \noindent
  This accomplishment would not have been possible without them.
  
  \bigskip \noindent
  Thank you.
  
  \bigskip \noindent
  Lukáš Forst
}

\keywordsCS{vyvažování a distribuce zátěže, optimalizační algoritmy, kombinatorická optimalizace, TASP, OptaPlanner, Kotlin, Ktor, Docker}
\keywordsEN{load balancing, optimization algorithms, combinatorial optimization, TASP, OptaPlanner Kotlin, Ktor, Docker}   


\abstractCS{ 
  \noindent
  Velké optimalizační úlohy nejsou jednoduché
  a pro nalezení řešení vyžadují velké množství výpočetního výkonu, stejně jako výpočetního času.
  Proto je výzvou je vyřešit paralelně a v co nejkratším možném čase.
  Optimalizační problémy nejsou homogenní skupina,
  ale popisují často kompletně odlišnou doménu a tedy mohou být velice různorodé.
  Problém exekuce optimalizačních programů
  a jejich výkonové náročnosti,
  lze řešit za použití inteligentního plánovače,
  který umožňuje naplánovat samotné spuštění úlohy v distribuovaném výpočetním prostředí.
 
  \noindent
  Tato práce se zabývá návrhem vyvažovače zátěže vyvinutého přímo pro optimalizační algoritmy,
  který minimalizuje plýtvání zdroji
  a zvyšuje výkon pomocí správného rozdělení využití mezi více instancí těchto algoritmů.
  Práce analyzuje nejmodernější řešení a technologie,
  které se používají k řešení takových problémů ve velkých infrastrukturách.
  Následně je navrženo nové řešení pro vyrovnávání zátěže specifické pro danou doménu.
  Práce obsahuje matematickou formalizaci problému optimalizace vyvažování zátěže
  a zároveň popisuje a řeší další nalezené problémy, jako je například predikce hyperbolických časových řad.
  Následně je v práci navržena architektura na bázi mikroslužeb pro systém vyvažování zátěže
  a také je daný systém kompletně realizován.
  Zároveň jsme navrhli a implementovali simulace a experimenty,
  které testují implementovaný vyvažovač zátěže.

  \noindent
  Závěrečná kapitole práce se zabývá vymezením nezbytných kroků pro úplnou produkcionalizaci navrženého systému vyvažování zátěže
  a také nastiňuje budoucí vývoj nástrojů a knihoven,
  které byly vyvinuty vedle primárního systému.
}

\abstractEN{
  \noindent
  Large-scale optimization problems are non-trivial to solve 
  and require a significant amount of computational resources as well as computational time to find a solution.
  The challenge is to solve not only a single task but a multitude of them in a parallel manner.
  Additionally, the tasks are non-homogeneous, 
  often describing a different problem. 
  This problem can be solved by designing 
  and intelligent scheduler able to schedule such tasks on a distributed computational platform.
  
  \noindent
  This work introduced the load balancer developed explicitly for the optimization algorithms,
  which should minimize resources wasting 
  and increase the performance using correct utilization distribution across the multiple instances of such algorithms.
  The thesis analyzes the state-of-the-art solutions and technologies,
  that are being used to solve load balancing problems in the large infrastructures.
  Subsequently, the new domain-specific load balancing solution is proposed.
  The thesis proposes the mathematical formalization of the load balancing optimization problem
  and the related challenge, such as hyperbola time series prediction.
  Subsequently, the thesis designs the microservices architecture for the load balancer 
  and also delivers the complete implementation of the proposed system.
  This thesis also proposed and implemented the simulations and experiments, 
  which evaluates the implemented load balancer.

  \noindent
  The final chapter of the thesis addresses out of scope steps for complete productionalization of the proposed load balancing system
  and also outlines the future development of the tools and libraries,
  that was developed alongside the primary system.
}

\begin{document}
    
\input{txt_c_intro.tex}

\input{txt_c_sota.tex}

\input{txt_c_problem_formalization.tex}

\input{txt_c_solution_design.tex}

\input{txt_c_implementation.tex}

\input{txt_c_experiments.tex}

\input{txt_c_conclusion.tex}

\bibliography{literature}

\appendix {
	\chapter{List of attachments}
	\begin{enumerate}
		\item CD with implementation described in chapter \ref{ch:implementation}
	\end{enumerate}
}
\end{document}