\section{Complete application algorithm}\label{sec:olb-algorithm}

The load balancing algorithm phases are divided into scheduling rounds.
Each scheduling round produces a new execution plan.
The execution plan is created for a specified scheduling horizon,
this horizon is the period,
which is limited by the time $t+1$ and the $t_{max}$, 
which defines the end of the execution plan
and where the $t$ is the time of execution.
In other words,
each scheduling round, executed in time $t$,
produces the scheduling plan with scheduling horizon from $t+1$ to $t_{max}$.
The scheduling horizon can be configured in the scheduling properties described in chapter \ref{ch:implementation}.
The length of the scheduling window can be configured as well. 

The formalized algorithm is then presented in the algorithm~\ref{alg:load-balancing-alg}.
The job related properties $D,T,P$ and $C$ are the very same variables defined in the section \ref{sec:formal-definition}.
The functions mentioned in the algorithm are described in the algorithm detailed description listed below the algorithm itself.
The function \textit{predict} is related to the step described in the step \ref{item:predictions-scheduling-step}
and the function \textit{schedule} to the step \ref{item:scheduling-step}.

\bigskip
\begin{algorithm}[H]
	\SetAlgoLined
	\Input{$Q$ - queue with jobs to schedule}
	$Q$: jobs queue\;
	$J$: set of jobs\;
	$P$: predictions\;
	$E$: execution plan\;
	\BlankLine
	$J \leftarrow$ $Q$.poll()\;
	$J'\leftarrow$ $J$.filter($job \rightarrow job.D > job.T \wedge job.P > job.C$)\;
	$J'' \leftarrow$ convert($J'$)\;
	$P \leftarrow$ predict($J''$)\;
	$E \leftarrow$ schedule($J''$, $P$)\;
	$E' \leftarrow$ convert($E$)\;
	\Result{$Q$ is empty}
	\Output{$E'$ - execution plan}
	\caption{Load balancing algorithm}
	\label{alg:load-balancing-alg}
\end{algorithm} 
\medskip
\begin{enumerate}
	\itemsep0pt
	\item \label{item:initial-scheduling-step} Poll jobs from the jobs queue.
	\begin{itemize}
		\item The incoming jobs from API are stored in the queue, 
		until the new scheduling round is engaged.
	\end{itemize}

	\item Analyze and filter jobs, that are not relevant for the current scheduling round.
	\begin{itemize}
		\item Filtered out are the jobs,
		 whose execution time exceeded the maximal possible execution time $D^{j}$
		 or whose cost has exceeded the maximal cost $P^{j}$.
	\end{itemize}

	\item \label{item:convert-to-scheduling-step} Convert received jobs into the inner data representation.
	\begin{itemize}
		\item The data must be converted from the data transfer objects structure
		to the OptaPlanner domain representation.
	\end{itemize}

	\item \label{item:predictions-scheduling-step} Create predictions based on the solution value of the jobs 
	and on the history of the load balancing decisions.
	\begin{itemize}
		\item The process of making the predictions is described in section \ref{sec:algorithm-value-prediction}.
	\end{itemize}

	\item \label{item:scheduling-step} Produce execution plan using OptaPlanner scheduling.
	\begin{itemize}
		\item The scheduling is limited to the amount of time defined as scheduling window.
		When the scheduling reaches the scheduling window (for example 60 seconds),
		it is stopped.
	\end{itemize}

	\item \label{item:convert-back-scheduling-step} Convert created plan into time schedule and send them back to the client.

	\item Engage the next scheduling round and go to the step \ref{item:initial-scheduling-step}.
\end{enumerate}	